---
title: Fort Collins, Colorado Employee Earnings
output: 
  github_document:
    html_preview: false
---

```{r cache=FALSE, echo=FALSE, message=F, warning=F}
knitr::read_chunk('main.R')
```
An analysis of
[2015 Fort Collins CO Employee Earnings](http://www.fcgov.com/openbook/?action=browse-salaries)

## Data overview
The data pulled from the Fort Collins Open Book website is available in the 
[data folder](data/raw). There were variables added to the dataset:

1. `serviceArea`: Using the 
[organizational chart](assets/fortCollinsOrganization.pdf), the the
`serviceArea` was added to organize the service units and departments into a
better structure.
2. `year`: To tidy the data for future use, a `year` variable of 2015 was added
to the data.
3. `hourlyWage`: [In 2015, The Associated Press reported](http://denver.cbslocal.com/2015/01/01/colorado-minimum-wage-rises-to-8-23-in-2015/)
that the minimum wage for Colorado was $8.23. Under the assumption that any
employee earning more than $17,506/year would be full time, we calculated their
possible hourly salary. Many employees are salaried, so this is more for 
informational purposes.
4. `status`: The status variable is used to help categorize employees into 
different buckets. Employees earning more than the assumed $8.23/hour would be 
labled as `full time` while anyone earning less would be `part time` or contract.


As of 2015, there are `r length(unique(salaries$serviceArea))` service areas
with a total of `r nrow(salaries)` employees.

### Data example
```{r echo = FALSE, message = FALSE, warning = FALSE}
knitr::kable(
  salaries %>%
    filter(serviceArea != "unknown") %>% 
    arrange(desc(earnings)) %>% 
    top_n(n = 10, earnings))
```

## Overview Table
```{r echo = FALSE, message = FALSE, warning = FALSE}
knitr::kable(salaries %>% 
  filter(serviceArea != "unknown") %>% 
  group_by(serviceArea) %>% 
  summarize(departmentCount = length(unique(department)),
            jobTitleCount = length(unique(jobTitle)),
            employeeCount = length(jobTitle),
            medianSalary = median(earnings)) %>% 
  arrange(desc(employeeCount)))
```